Wigglyhedra

### 2024-07-12 draft/wigglyhedra.tex
More updates (some in progress).

  We recall that the procedure to build up \(X\) from \(\alpha\) explicitly describes the differential.
  We analyse the possible non-zero components of the differential, checking how each interacts with the level.
  The only two possibilities are as follows:
  \begin{enumerate}
  \item a map \(P_i\langle g \rangle[h] \to P_{i\pm1}\langle g+1 \rangle[h-1]\), namely multiplication on the right by the path \((i,i\pm1)\), which does not change the level; or
  \item a map \(P_i\langle g \rangle[h] \to P_{i}\langle g+2 \rangle[h-1]\), namely multiplication on the right by the ``loop'' path \((i,i\pm1, i)\), which raises the level by one.
  \end{enumerate}
  The first possibility appears each time the curve wiggles either above the \(i\)th point or below the \(i-1\)th point.
  The second possibility appears each time the curve makes a \(U\)-turn around the \(i\)th or \(i-1\)th point.

@@

  The object \(X\) is in the heart up to shift if and only if it is concentrated in a single level \(j\).
  Truncating to level \(j\) as described above, we do not delete any of the terms of \(X\).
  This can happen if and only if the differential of \(X\) contains no loop components,
  which appear in the complex if and only if \(\alpha\) makes a U-turn around one of the \(n+2\) points.
  Thus \(X\) is in the heart up to shift if and only if \(\alpha\) is  \(x\)-monotone.
  This proves the first part of the proposition.

  More generally, consider the indecomposable pieces of the level-\(j\) truncations of \(X\) for various~\(j\), up to shift.
  As explained above, in terms of the curve \(\alpha\), this corresponds to breaking up the curve~\(\alpha\) into \(x\)-monotone portions.
  More precisely, we begin at an endpoint \(p_0\) of \(\alpha\) and follow the curve until it either ends at a point \(p_1\), or stops being \(x\)-monotone by bending around a point \(p_1\).
  In this case, we take the portion of \(\alpha\) between \(p_0\) and \(p_1\), calling it \(\alpha_1\), and setting its endpoints to be \(p_0\) and \(p_1\) respectively.
  Note that \(p_0 \neq p_1\) because \(\alpha\) is \(x\)-monotone between \(p_0\) and \(p_1\).
  We continue with the remainder of \(\alpha\) starting at \(p_1\), greedily taking the maximal \(x\)-monotone portion each time, to obtain pieces \(\alpha_1,\alpha_2, \ldots, \alpha_k\), which are all wiggly arcs.
  These wiggly arcs correspond to complexes \(X_1, \ldots, X_k\), which are spherical objects of \(\mathcal{A}_m\), up to shift.
  It is clear from the construction of the associated complex that these objects, up to shift, are exactly the indecomposable pieces of the cohomology truncations \(H^j(X)\) for various \(j\).
  This argument proves the second part of the proposition.
\end{proof}

@@

  We have four possible combinations:
  \begin{enumerate}
  \item \(i' = j\) is an endpoint such that \(\alpha\) goes left starting at \(i'\), and \(i' = j\) is an endpoint such that \(\alpha\) goes right starting at \(j\);
  \item \(i' = j\) is an endpoint such that \(\alpha\) goes left starting at \(i'\), and \(\alpha\) loops around the left of the point \(j\);
  \item \(\alpha\) loops around the right of the point \(i' = j\), and \(i' = j\) is an endpoint such that \(\alpha\) goes right starting at \(j\); 
  \item \(\alpha\) loops around the right of the point \(i' = j\), and \(\alpha\) loops around the left of the point \(j\).
  \end{enumerate}
  It is clear that none of these possibilities can occur within the curve \(\alpha\), and hence \(\beta\) and \(\beta'\) must be pointed.
  Therefore \(T = \{\alpha_1, \ldots, \alpha_k\}\) is a subset of a wiggly pseudotriangulation.

  Now suppose for contradiction that \(T\) contains both external edges.
  Then some portion of \(\alpha\) curves around each of the extreme endpoints \(0\) and \(n+1\).
  Consider the outermost strands of \(\alpha\) that curve around \(0\) and \(n+1\) respectively.
  These must then connect to each other, resulting in \(\alpha\) having a separate component that winds around the outside of the \(n+2\) points.
  This is not possible because \(\alpha\) is a connected curve with two endpoints.
  Therefore \(T\) cannot contain both external edges.
\end{proof}

@@

\begin{proof}
  The proof is once again a consequence of analysing the extensions explicitly via the complexes from the Khovanov--Seidel reconstruction.
  We do not reproduce the proof here, as the argument is contained in the proof of~\cite[Thm.~5.5]{tho:06}.
  (See also~\cite[Fig.~4]{tho:06}.)
\end{proof}
Note that if we write down the relations in the Grothendieck group of our category that correspond to the extensions from each case of~\cref{prop:extensions}, then we obtain precisely the linear dependences that appear between the \(\b{g}\)-vectors in~\cref{lem:linearDependences}.
More precisely, we have that
\begin{itemize}
\item if \(\alpha\) and \(\alpha'\) are not pointed, then
  \([Y] = [X] + [X'] = [Y']\),
  from which we see \([X] + [X'] = ([Y] + [Y'])/2\);
\item if \(\alpha\) and \(\alpha'\) are crossing, then
  \([X] + [X'] = [Y] + [Y'].\)
\end{itemize}
This suggests that there should be a way to construct the \(\b{g}\)-vectors categorically in such a way that the \(\b{g}\)-vectors also map to the Grothendieck group of the category.
At the moment, we do not have a concise construction of the \(\b{g}\)-vectors purely categorically, and it would be interesting to find this.

>>>>

  We recall that the procedure to build up \(X\) from \(\alpha\) explicitly describes the differential.
  We analyse the possible non-zero components of the differential, checking how each interacts with the level.
  The only two possibilities are as follows:
  \begin{enumerate}
  \item a map \(P_i\langle g \rangle[h] \to P_{i\pm1}\langle g+1 \rangle[h-1]\), namely multiplication on the right by the path \((i,i\pm1)\), which does not change the level; or
  \item a map \(P_i\langle g \rangle[h] \to P_{i}\langle g+2 \rangle[h-1]\), namely multiplication on the right by the ``loop'' path \((i,i\pm1, i)\), which raises the level by one.
  \end{enumerate}
  The first possibility appears each time the curve wiggles either above the \(i\)th point or below the \((i-1)\)th point.
  The second possibility appears each time the curve makes a U-turn around the right of \(i\)th or the left of the \((i-1)\)th point.

@@

  The object \(X\) is in the heart up to shift if and only if it is concentrated in a single level \(j\).
  Truncating to level \(j\) as described above, we do not delete any of the terms of \(X\).
  This can happen if and only if the differential of \(X\) contains no ``loop'' components,
  which appear in the complex if and only if \(\alpha\) makes a U-turn around one of the \(n+2\) points.
  Thus \(X\) is in the heart up to shift if and only if \(\alpha\) is  \(x\)-monotone.
  This proves the first part of the proposition.

  More generally, consider the indecomposable pieces of the level-\(j\) truncations of \(X\) for various~\(j\), up to shift.
  As explained above, in terms of the curve \(\alpha\), this corresponds to breaking up the curve~\(\alpha\) into \(x\)-monotone portions.
  More precisely, we begin at an endpoint \(p_0\) of \(\alpha\) and follow the curve until it either ends at a point \(p_1\), or stops being \(x\)-monotone by making a U-turn around a point \(p_1\).
  In this case, we take the portion of \(\alpha\) between \(p_0\) and \(p_1\), calling it \(\alpha_1\), and setting its endpoints to be \(p_0\) and \(p_1\) respectively.
  Note that \(p_0 \neq p_1\) because \(\alpha\) is \(x\)-monotone between \(p_0\) and \(p_1\).
  We continue with the remainder of \(\alpha\) starting at \(p_1\), greedily taking the maximal \(x\)-monotone portion each time, to obtain pieces \(\alpha_1,\alpha_2, \ldots, \alpha_k\), which are all wiggly arcs.
  These wiggly arcs correspond to complexes \(X_1, \ldots, X_k\), which are spherical objects of \(\mathcal{A}_m\), up to shift.
  It is clear from the construction of the associated complex that these objects, up to shift, are exactly the indecomposable pieces of the cohomology truncations \(H^j(X)\) for various \(j\).
  This argument proves the second part of the proposition.
\end{proof}

@@

  We have four possible combinations:
  \begin{enumerate}
  \item \(i' = j\) is an endpoint such that \(\alpha\) travels left starting at \(i'\), and \(i' = j\) is an endpoint such that \(\alpha\) travels right starting at \(j\);
  \item \(i' = j\) is an endpoint such that \(\alpha\) travels left starting at \(i'\), and \(\alpha\) loops around the left of the point \(j\);
  \item \(\alpha\) loops around the right of the point \(i' = j\), and \(i' = j\) is an endpoint such that \(\alpha\) travels right starting at \(j\); 
  \item \(\alpha\) loops around the right of the point \(i' = j\), and \(\alpha\) loops around the left of the point \(j\).
  \end{enumerate}
  It is clear that none of these possibilities can occur within the curve \(\alpha\), and hence \(\beta\) and \(\beta'\) must be pointed.
  Therefore \(T = \{\alpha_1, \ldots, \alpha_k\}\) is a subset of a wiggly pseudotriangulation.

  Now suppose for contradiction that \(T\) contains both external edges.
  Then some portion of \(\alpha\) curves around each of the extreme endpoints \(0\) and \(n+1\).
  Consider the outermost strands of \(\alpha\) that curve around \(0\) and \(n+1\) respectively.
  These must then connect to each other, resulting in \(\alpha\) having a separate closed component that winds around the outside of the \(n+2\) points.
  This is not possible because \(\alpha\) is a connected curve with two endpoints.
  Therefore \(T\) cannot contain both external edges.
\end{proof}

@@

\begin{proof}
  The proof is once again a consequence of analysing the extensions explicitly via the complexes from the Khovanov--Seidel reconstruction.
  We do not reproduce the proof here, as the argument is contained in the proof of~\cite[Thm.~5.5]{tho:06}.
  (See also~\cite[Fig.~4]{tho:06}.)
\end{proof}
Note that if we write down the relations in the Grothendieck group of our category that correspond to the extensions from each case of~\cref{prop:extensions}, then we obtain precisely the linear dependences that appear between the \(\b{g}\)-vectors in~\cref{lem:linearDependences}.
More precisely, we have that
\begin{itemize}
\item if \(\alpha\) and \(\alpha'\) are not pointed, then
  \([Y] = [X] + [X'] = [Y']\),
  from which we see \([X] + [X'] = ([Y] + [Y'])/2\);
\item if \(\alpha\) and \(\alpha'\) are crossing, then
  \([X] + [X'] = [Y] + [Y'].\)
\end{itemize}
This suggests that there should be a way to construct the \(\b{g}\)-vectors categorically in such a way that the \(\b{g}\)-vectors also map to the Grothendieck group of the category.
The \(\b{g}\)-vector of an object \(X\) corresponding to a wiggly arc \(\alpha\) can, of course, be reconstructed from \(\alpha\).
However, at the moment we do not have a purely categorical construction of the \(\b{g}\)-vector of \(X\) that does not refer to \(\alpha\), and it would be interesting to find this.
